Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-09-22 manuscript/usses_water_ms.Rmd
rewrite discussion after PBA comments

  **Discussion.** Despite some evidence that ecosystem functional response became less sensitive under sustained wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable. Differences in ecosystem functional responses across treatments were not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture. These species may be successful precisely because they have evolved life history strategies which buffer them against precipitation variability.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. 
$\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. 
Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of VWC, binary "drought" value times VWC, and binary "irrigation" value times VWC. 
Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. 
We use our model to test the following hypotheses:

@@

Ecosystem response to precipitation extremes depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
A lingering question is how the time scales of ecosystem response and community change vary among ecosystems. Precipitation manipulation experiments can help answer this question, especially if they push water availability outside the historical range of variability for long periods.

The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
Neither ecosystem functional response nor community composition changed under chronic drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig. 3A, Fig. 4), representing the top right SCENARIO in Fig. 1.
The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a saturating response for the irrigation treatment (Fig. 3).
<!--- Can we really call it saturating if it is concave up? More accurate to say weaker response than control? --->
However, the response of ANPP to soil moisture in all treatments was consistently weak (Table 1, Fig. A2-2) and within-year variation of ANPP was large (Fig. 3B).
<!---Last sentence leaves me hanging...what's your point? Did it the response change or not? Statistically but not ecologically significant?--->

The similarity of ecosystem functional response (Fig. 3) and community composition (Fig. 5) between drought and control treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
Indeed, we expected ecosystem functional response, community composition, or both to change under the drought treatment, landing us in any box of Fig. 1 \emph{except} the top left.
Why did our drought treatment fail to induce ecosystem or community responses?
We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
We first discuss the potential limitations of our study, and then discuss the biological explanation.

First, it could be that our drought manipulation was not large enough to induce a response.
That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
We cannot definitively rule out this possibility, but we have reason to believe our drought treatment *should* have been large enough.
Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig. A3-3).
The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3).

Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig. A2-2).
<!---so why didn't we do this? sample size limitations, power? --->

Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Fig. 3).
Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
In other words, species follow the same conservative strategy every year, designed to minimize response to environmental conditions.
The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes in $\boldsymbol{\beta}$ in Eq. 1 overlapping zero).

@@

The weaker soil moisture-ANPP relationship we observed for the irrigation treatment has three non-exclusive explanations.
First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR. YOU COULD CITE THIS AS UNPUBLISHED DATA}

Second, the less positive slope could be an indicator of bet hedging in this community.
Avoiding high biomass and seed production in wet years could benefit species in the long run if climate conditions are variable.
\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.

In conclusion, our results suggest the species in our focal plant community are tolerant of drought conditions and unresponsive to wet conditions. \pba{no, you can't be a bet hedger in wet years but not in dry years...}
Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016]. For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.

>>>>

  **Discussion.** Despite some evidence that ecosystem functional response became less sensitive under sustained wet conditions, the response of ANPP to soil moisture was consistently weak and community composition was stable. The similarity of ecosystem functional responses across treatments was not related to compensatory shifts at the plant community level, but instead may reflect the insensitivity of the dominant species to soil moisture. These species may be successful precisely because they have evolved life history strategies which buffer them against precipitation variability.
header-includes: >
  \usepackage{setspace}
  \usepackage{todonotes}
  \usepackage{rotating}
  \usepackage{color, soul}
  \usepackage{sectsty}
  \usepackage{bm,mathrsfs}
bibliography: /Users/atredenn/Dropbox/Bibliography/usses_water.bib
output: 
  rticles::peerj_article:
    number_sections: false
    base_format: rmarkdown::pdf_document # bookdown::pdf_document2 # for using \@ref()
---

@@

\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. 
$\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. 
Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of VWC, binary "drought" value times VWC, and binary "irrigation" value times VWC. 
Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. 
We use our model to test two statistical hypotheses based on the questions in our Introduction:

@@

Ecosystem response to a new precipitation regime depends on the physiological responses of constituent species and the rate at which community composition shifts to favor species better able to take advantage of, or cope with, new resource levels [@Smith2009].
Previous work has shown that community compositional shifts can be both rapid, on the order of years [@Hoover2014], and slow, on order of decades [@Knapp2012; @Wilcox2016].
A lingering question is how the time scales of ecosystem response and community change vary among ecosystems. 
Precipitation manipulation experiments can help answer this question, especially if they push water availability outside the historical range of variability for long periods.

The results of our five year experiment in a sagebrush steppe conform to two of our four predictions, depending on treatment.
Ecosystem functional response and community composition did not change under chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in Fig. 1.
Ecosystem functional response under chronic irrigation was different from the control treatment, but community composition remained unchanged (Fig. 3A, Fig. 4), representing the top right scenario in Fig. 1.
The altered ecosystem functional response under irrigation matched our expectations because we found evidence for a weaker response for the irrigation treatment (Fig. 3).
Although the difference between the control and irrigation slopes is likely different from zero (Fig. 3A), the posterior distribution of the slopes (not their difference) are similar and overlap broadly (Table1, Fig. A2-2).
Thus, the statistical difference between the slopes is not ecologically meaningful.
We therefore conclude that ecosystem functional response is consistent and weak across all precipitation treatments.

The similarity of ecosystem functional response (Fig. 3) and community composition (Fig. 5) among treatments is surprising because grasslands generally, and sagebrush steppe specifically, are considered water-limited systems.
Indeed, we expected ecosystem functional response, community composition, or both to change under our treatments.
Why did our treatments fail to induce ecosystem or community responses?
We can think of three reasons; two are limitations of our study, and one is the life history traits of the species in our focal communities.
We first discuss the potential limitations of our study, and then discuss the biological explanation.

First, it could be that our precipitation manipulations were not large enough to induce a response.
That is, a 50\% decrease in any given year may not be abnormal given our site's historical range of variability [@Knapp2017].
We cannot definitively rule out this possibility, but we have reason to believe our treatments *should* have been large enough.
Using the methods described by @Lemoine2016, we calculated the percent reduction and increase of mean growing season precipitation necessary to reach the 1\% and 99\% extremes of the historical precipitation regime at our site (Fig. A3-3).
The 1\% quantile of precipitation at our site is 110 mm, a 47% reduction from the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean growing season precipitation (Appendix 3).
Thus, our drought treatment represented extreme precipitation amounts, especially in years where ambient precipitation was below average (Fig. 2A).
The irrigation treatment may not qualify as extreme, yet that is the treatment where we did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not only the cumulative March-June soil moisture covariate we included in our statistical model.
For example, temperature can impact ANPP directly [@Epstein1997] and by exacerbating the effects of soil moisture [@DeBoeck2011].
Measurements of soil moisture likely contain a signal of temperature, through its impact on evaporation and infiltration, but the measurements will not capture the direct effect of temperature on metabolic and physiological processes.
We also did not redistribute snow across our treatments in the winter, and snow melt may spur early spring growth.
Failure to account for potentially important covariates could explain the within-year spread of ANPP (Fig 2C, Fig. 3B) and the resulting uncertain relationship we observed between soil moisture and ANPP across all treatments (Table 1, Fig. A2-2).
Unfortunately, we did not collect data on additional factors because our focus was on precipitation.

Third, the life history traits of the dominant species in our study ecosystem may explain the consistently positive, but weak and uncertain, effect of soil moisture on ANPP (Table 1, Fig. 3).
Species that live in variable environments, such as cold deserts, must have strategies to ensure long-term success as conditions vary.
One strategy is bet hedging, where species forego short-term gains to reduce the variance of long-term success [@Seger1987].
In other words, species follow the same conservative strategy every year, designed to minimize response to environmental conditions.
The dry and variable environment of the sagebrush steppe has likely selected for bet hedging species that can maintain function at low water availability and have weak responses to high water availability.
In so doing, the dominant species in our ecosystem avoid "boom and bust" cycles, which corresponds to the weak effect of soil moisture on ANPP (i.e., the Bayesian credible intervals for the slopes in $\boldsymbol{\beta}$ in Eq. 1 overlapping zero).

@@

<!---
The weaker soil moisture-ANPP relationship we observed for the irrigation treatment has three non-exclusive explanations.
First, species' biomass production may have hit physiological thresholds and/or become limited by other factors, such as nitrogen availability [@LeBauer2008].
\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR. YOU COULD CITE THIS AS UNPUBLISHED DATA}

Second, the less positive slope could be an indicator of bet hedging in this community.
Avoiding high biomass and seed production in wet years could benefit species in the long run if climate conditions are variable.
\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
Third, the less positive slope may be a statistical artifact due the variability of ANPP within years and our low sample size.
--->

In conclusion, our results suggest the species in our focal plant community are insensitive to to changes in precipitation regime. 
Such insensitivity can buffer species against precipitation variability in this semi-arid ecosystem, making them successful in the long run.
Longer, chronic precipitation alteration might reveal plant community shifts that we did not observe [e.g., @Wilcox2016]. 
For example, a long-term increase in water availability could allow species that do not bet hedge to gain prominence and dominate the ecosystem functional response.
Our results suggest compositional shifts would have the largest impact at high rainfall because the current community maintained consistent ecosystem functional response at very low water availability.

### 2017-09-22 manuscript/usses_water_ms.tex
rewrite discussion after PBA comments

\textbf{Discussion.} Despite some evidence that ecosystem functional
response became less sensitive under sustained wet conditions, the
response of ANPP to soil moisture was consistently weak and community
composition was stable. Differences in ecosystem functional responses
across treatments were not related to compensatory shifts at the plant
community level, but instead may reflect the insensitivity of the
dominant species to soil moisture. These species may be successful
precisely because they have evolved life history strategies which buffer
them against precipitation variability.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \emph{i}, which is associated with plot
\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
vector of coefficients for the fixed effects in the design matrix
\(\textbf{X}\). Each row of the design matrix represents a single
observation (\(\textbf{x}_i\)) and is a vector with the following
elements: 1 for the intercept, a binary 0 or 1 if the treatment is
``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
scaled value of VWC, binary ``drought'' value times VWC, and binary
``irrigation'' value times VWC. Thus, our model treats ``control''
observations as the main treatment and then estimates intercept and
slope offsets for the ``drought'' and ``irrigation'' treatments. We use
our model to test the following hypotheses:

@@

Ecosystem response to precipitation extremes depends on the
physiological responses of constituent species and the rate at which
community composition shifts to favor species better able to take
advantage of, or cope with, new resource levels \citep{Smith2009}.
Previous work has shown that community compositional shifts can be both
rapid, on the order of years \citep{Hoover2014}, and slow, on order of
decades \citep{Knapp2012, Wilcox2016}. A lingering question is how the
time scales of ecosystem response and community change vary among
ecosystems. Precipitation manipulation experiments can help answer this
question, especially if they push water availability outside the
historical range of variability for long periods.

The results of our five year experiment in a sagebrush steppe conform to
two of our four predictions, depending on treatment. Neither ecosystem
functional response nor community composition changed under chronic
drought (Fig. 3A, Fig. 4), representin the top left SCENARIO in Fig. 1.
Ecosystem functional response under chronic irrigation was different
from the control treatment, but community composition remained unchanged
(Fig. 3A, Fig. 4), representing the top right SCENARIO in Fig. 1. The
altered ecosystem functional response under irrigation matched our
expectations because we found evidence for a saturating response for the
irrigation treatment (Fig. 3). However, the response of ANPP to soil
moisture in all treatments was consistently weak (Table 1, Fig. A2-2)
and within-year variation of ANPP was large (Fig. 3B).

The similarity of ecosystem functional response (Fig. 3) and community
composition (Fig. 5) between drought and control treatments is
surprising because grasslands generally, and sagebrush steppe
specifically, are considered water-limited systems. Indeed, we expected
ecosystem functional response, community composition, or both to change
under the drought treatment, landing us in any box of Fig. 1
\emph{except} the top left. Why did our drought treatment fail to induce
ecosystem or community responses? We can think of three reasons; two are
limitations of our study, and one is the life history traits of the
species in our focal communities. We first discuss the potential
limitations of our study, and then discuss the biological explanation.

First, it could be that our drought manipulation was not large enough to
induce a response. That is, a 50\% decrease in any given year may not be
abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our drought treatment \emph{should} have been
large enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A3-3). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 3). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3).

Second, ANPP at our site may be influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. For example, temperature can impact ANPP directly
\citep{Epstein1997} and by exacerbating the effects of soil moisture
\citep{DeBoeck2011}. Measurements of soil moisture likely contain a
signal of temperature, through its impact on evaporation and
infiltration, but the measurements will not capture the direct effect of
temperature on metabolic and physiological processes. We also did not
redistribute snow across our treatments in the winter, and snow melt may
spur early spring growth. Failure to account for potentially important
covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
3B) and the resulting uncertain relationship we observed between soil
moisture and ANPP across all treatments (Table 1, Fig. A2-2).

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistently positive, but weak and uncertain,
effect of soil moisture on ANPP (Fig. 3). Species that live in variable
environments, such as cold deserts, must have strategies to ensure
long-term success as conditions vary. One strategy is bet hedging, where
species forego short-term gains to reduce the variance of long-term
success \citep{Seger1987}. In other words, species follow the same
conservative strategy every year, designed to minimize response to
environmental conditions. The dry and variable environment of the
sagebrush steppe has likely selected for bet hedging species that can
maintain function at low water availability and have weak responses to
high water availability. In so doing, the dominant species in our
ecosystem avoid ``boom and bust'' cycles, which corresponds to the weak
effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
for the slopes in \(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).

@@

The weaker soil moisture-ANPP relationship we observed for the
irrigation treatment has three non-exclusive explanations. First,
species' biomass production may have hit physiological thresholds and/or
become limited by other factors, such as nitrogen availability
\citep{LeBauer2008}.
\pba{PETER: DO WE KNOW ANYTHING ABOUT NITROGEN AT USSES? WANT TO ADD A SENTENCE HERE ABOUT HOW LIKELY OR UNLIKEL THIS IS...NUTNET SHOWS POSITIVE EFFECTS OF N*P ON PRODUCTION, EFFECT OF N ALONE NOT AS CLEAR. YOU COULD CITE THIS AS UNPUBLISHED DATA}

Second, the less positive slope could be an indicator of bet hedging in
this community. Avoiding high biomass and seed production in wet years
could benefit species in the long run if climate conditions are
variable.
\pba{but why would this have cause a change relative to controls? this should also be true for controls...}
Third, the less positive slope may be a statistical artifact due the
variability of ANPP within years and our low sample size.

In conclusion, our results suggest the species in our focal plant
community are tolerant of drought conditions and unresponsive to wet
conditions.
\pba{no, you can't be a bet hedger in wet years but not in dry years...}
Longer, chronic precipitation alteration might reveal plant community
shifts that we did not observe \citep[e.g.,][]{Wilcox2016}. For example,
a long-term increase in water availability could allow species that do
not bet hedge to gain prominence and dominate the ecosystem functional
response. Our results suggest compositional shifts would have the
largest impact at high rainfall because the current community maintained
consistent ecosystem functional response at very low water availability.

>>>>

\textbf{Discussion.} Despite some evidence that ecosystem functional
response became less sensitive under sustained wet conditions, the
response of ANPP to soil moisture was consistently weak and community
composition was stable. The similarity of ecosystem functional responses
across treatments was not related to compensatory shifts at the plant
community level, but instead may reflect the insensitivity of the
dominant species to soil moisture. These species may be successful
precisely because they have evolved life history strategies which buffer
them against precipitation variability.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \emph{i}, which is associated with plot
\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
vector of coefficients for the fixed effects in the design matrix
\(\textbf{X}\). Each row of the design matrix represents a single
observation (\(\textbf{x}_i\)) and is a vector with the following
elements: 1 for the intercept, a binary 0 or 1 if the treatment is
``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
scaled value of VWC, binary ``drought'' value times VWC, and binary
``irrigation'' value times VWC. Thus, our model treats ``control''
observations as the main treatment and then estimates intercept and
slope offsets for the ``drought'' and ``irrigation'' treatments. We use
our model to test two statistical hypotheses based on the questions in
our Introduction:

@@

Ecosystem response to a new precipitation regime depends on the
physiological responses of constituent species and the rate at which
community composition shifts to favor species better able to take
advantage of, or cope with, new resource levels \citep{Smith2009}.
Previous work has shown that community compositional shifts can be both
rapid, on the order of years \citep{Hoover2014}, and slow, on order of
decades \citep{Knapp2012, Wilcox2016}. A lingering question is how the
time scales of ecosystem response and community change vary among
ecosystems. Precipitation manipulation experiments can help answer this
question, especially if they push water availability outside the
historical range of variability for long periods.

The results of our five year experiment in a sagebrush steppe conform to
two of our four predictions, depending on treatment. Ecosystem
functional response and community composition did not change under
chronic drought (Fig. 3A, Fig. 4), representing the top left scenario in
Fig. 1. Ecosystem functional response under chronic irrigation was
different from the control treatment, but community composition remained
unchanged (Fig. 3A, Fig. 4), representing the top right scenario in Fig.
1. The altered ecosystem functional response under irrigation matched
our expectations because we found evidence for a weaker response for the
irrigation treatment (Fig. 3). Although the difference between the
control and irrigation slopes is likely different from zero (Fig. 3A),
the posterior distribution of the slopes (not their difference) are
similar and overlap broadly (Table1, Fig. A2-2). Thus, the statistical
difference between the slopes is not ecologically meaningful. We
therefore conclude that ecosystem functional response is consistent and
weak across all precipitation treatments.

The similarity of ecosystem functional response (Fig. 3) and community
composition (Fig. 5) among treatments is surprising because grasslands
generally, and sagebrush steppe specifically, are considered
water-limited systems. Indeed, we expected ecosystem functional
response, community composition, or both to change under our treatments.
Why did our treatments fail to induce ecosystem or community responses?
We can think of three reasons; two are limitations of our study, and one
is the life history traits of the species in our focal communities. We
first discuss the potential limitations of our study, and then discuss
the biological explanation.

First, it could be that our precipitation manipulations were not large
enough to induce a response. That is, a 50\% decrease in any given year
may not be abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our treatments \emph{should} have been large
enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A3-3). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 3). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3A).

Second, ANPP at our site may be influenced by additional factors, not
only the cumulative March-June soil moisture covariate we included in
our statistical model. For example, temperature can impact ANPP directly
\citep{Epstein1997} and by exacerbating the effects of soil moisture
\citep{DeBoeck2011}. Measurements of soil moisture likely contain a
signal of temperature, through its impact on evaporation and
infiltration, but the measurements will not capture the direct effect of
temperature on metabolic and physiological processes. We also did not
redistribute snow across our treatments in the winter, and snow melt may
spur early spring growth. Failure to account for potentially important
covariates could explain the within-year spread of ANPP (Fig 2C, Fig.
3B) and the resulting uncertain relationship we observed between soil
moisture and ANPP across all treatments (Table 1, Fig. A2-2).
Unfortunately, we did not collect data on additional factors because our
focus was on precipitation.

Third, the life history traits of the dominant species in our study
ecosystem may explain the consistently positive, but weak and uncertain,
effect of soil moisture on ANPP (Table 1, Fig. 3). Species that live in
variable environments, such as cold deserts, must have strategies to
ensure long-term success as conditions vary. One strategy is bet
hedging, where species forego short-term gains to reduce the variance of
long-term success \citep{Seger1987}. In other words, species follow the
same conservative strategy every year, designed to minimize response to
environmental conditions. The dry and variable environment of the
sagebrush steppe has likely selected for bet hedging species that can
maintain function at low water availability and have weak responses to
high water availability. In so doing, the dominant species in our
ecosystem avoid ``boom and bust'' cycles, which corresponds to the weak
effect of soil moisture on ANPP (i.e., the Bayesian credible intervals
for the slopes in \(\boldsymbol{\beta}\) in Eq. 1 overlapping zero).

@@

In conclusion, our results suggest the species in our focal plant
community are insensitive to to changes in precipitation regime. Such
insensitivity can buffer species against precipitation variability in
this semi-arid ecosystem, making them successful in the long run.
Longer, chronic precipitation alteration might reveal plant community
shifts that we did not observe \citep[e.g.,][]{Wilcox2016}. For example,
a long-term increase in water availability could allow species that do
not bet hedge to gain prominence and dominate the ecosystem functional
response. Our results suggest compositional shifts would have the
largest impact at high rainfall because the current community maintained
consistent ecosystem functional response at very low water availability.
